Real-time resolution of short-read assembly graph using ONT long reads

### 2020-03-02 paper.tex
plos

\title{Real-time resolving assembly graph by ONT long reads data}
\shorttitle{Real-time Scaffolding with Assembly graph}

%Authors
\author[1,$\ast$]{Son Hoang Nguyen}
\author[2]{Minh Duc Cao}
\author[1,$\ast$]{Lachlan Coin}

%Affiliation
\affil[1]{Institute for Molecular Bioscience, the University of Queensland, 
St Lucia, Brisbane, QLD 4072 Australia}
\affil[2]{4catalyzer}
\correspondingauthor{
\textsuperscript{$\ast$}
To whom correspondence should be addressed. 
E-mails: l.coin@imb.uq.edu.au,s.hoangnguyen@imb.uq.edu.au
}

@@

%Set abstract (see below)
\abstract{
A real-time assembly pipeline for Oxford Nanopore Technology (ONT) data is important for either saving sequencing resources or reducing turnaround time for data analyses. The previous approach from \npscarf{} provided a fast-response streaming algorithm for such task but was relatively prone to mis-assemblies compared to other graph-based methods. 
Here we present \npgraph{}, a real-time hybrid assembly software using the assembly graph instead of the separated pre-assembly contigs. It is able to produce more complete genome assembly by resolving the path finding problem on the assembly graph using long reads as the traversing guide. 
The application on synthetic and real data set of isolate genomes show improved accuracy while still maintaining the required computational cost extremely low. 
%On top of that, we apply \npgraph{} on mock community in the very first application to improve metagenomics assembly in real-time with nanopore data. 
With \npscarf{}, the built-in graphical user interface (GUI) can provide users a comprehensive look-and-feel of the whole assembly process. The tool and source code is available at \url{https://github.com/hsnguyen/assembly}.
} 
%Set keywords
\keywords{hybrid assembly, assembly graph, real-time analysis, nanopore sequencing}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}
% \begin{itemize}
% \item The importance of having a real-time assembler for bacterial and small genomes: e.g. npscarf 
% \item The difficulty to resolve assembly graph in real-time.
% \item Available hybrid assembly methods using long reads: unicycler, ...
% \item an online algorithm to resolve assembly graph by long reads is needed, ...
% \item Requirements
%     \begin{itemize}
%     \item Support streaming: efficient with high yield (armortized analysis)
%     \item Statistically robust: online modification of the dynamic graph, back tracking edge deletion...
%     \item Standardisation : output GFA in real-time
%     \end{itemize}
% \end{itemize}
Streaming assembly methods had been proven to be useful in saving time and resources compared to the conventional batch algorithms with examples included \EG{} $\mathtt{Faucet}$~\cite{Rozov2017faucet} and \npscarf{}~\cite{Cao2017scaffolding}. The first method allows the assembly graph to be constructed incrementally as long as reads are retrieved and processed. This practice is helpful dealing with huge short-read data set because it can significantly reduce the local storage for the reads, as well as save time for a De Bruijn graph (DBG) construction while waiting for the data being retrieved.
\npscarf{}, on the other hand, is a hybrid assembler working on a pre-assembly set of short-read assembly contigs. It functions by scaffolding the contigs using nanopore sequencing which is well-known by the real-time property. The completion of genome assembly in parallel with the sequencing run provides explicit benefits in term of resource control and turn-around time for analysis~\cite{Cao2017scaffolding}.

% Why hybrid???
Hybrd approaches are still common practice in genome assembly and data analyses when Illumina sequencing remains the most favoured option in terms of cost and accuracy to date.
On the other hand, the third-generation sequencing methods such as Pacbio or Oxford Nanopore Technology are well-known for the ability to produce much longer reads that can further complete the Illumina assembly.
As the consequence, it is rational to combine two sources of data together in a hybrid method that can offer accurate and comptete genomes at the same time.
\npscarf{}, following that philosophy, had been developed and deployed on real microbial genomes.

However, due to the greedy bridging approach of the contigs-based streaming algorithm, \npscarf{}`s results might suffer from mis-assemblies~\cite{Wick2017unicycler,Giordano2017}. 
A default setting were optimzed for microbial genomes input but cannot fit for all data from various experiments in practice.
Also, the gap filling step has to rely on the low quality nanopore reads thus the accuracy of the final assembly is affected as well. 
To tackle the quality issue while maintaining its streaming feature, a bridging method by assembly graph traversing is proposed. 
In which, after the construction of a compact DBG assembly graph, the next step is to traverse the graph, resolve the repeats and identify the longest possible un-branched paths that would represents contigs for the final assembly.

Hybrid assembler using nanopore data to resolve the graph has been implemented in $\mathtt{hybridSPAdes}$ \cite{AntipovKM2015} or \unicycler{} \cite{Wick2017unicycler}. 
In general, the available tools employ batch-mode algorithms on the whole long-read data set to generate the final genome assembly. 
In which, the \spades{} hybrid assembly module, from its first step, exhaustively looks for the most likely paths (with mininum edit distance) on the graph for each of the long read given but only ones supported by at least two reads are attained. In the next step, these paths will be subjected to a decision-rule algorithm, namely $\mathtt{exSPAnder}$~\cite{Prjibelski2014}, for repeat resolution by step-by-step expansion, before output the final assembly.
On the other hand, \unicycler{}'s hybrid assembler will initially generate a consensus long read for each of the bridge from the batch data. 
The higher quality consensus reads are used to align with the assembly graph to find the best paths bridging pairs of anchored contigs.
While the latter method employs the completeness of the data  set from the very beginning for a consensus step, the former only iterates over the batch of possible paths and relies on a scoring system for the final decision of graph traversal. For that reason, the first direction is more suitable for a real-time pipeline.

Nonetheless, the challenge to adapt this approach into a real-time mechanism is obvious, mainly from building a progressive implementation for path-finding and graph reducing module which are essential for a streaming graph assembler. 
To achieve that, we apply a modified DFS (depth-first search) mechanism and a dynamic voting algorithm into an on-the-fly graph resolver.
The method is implemented in \npgraph{}, a user-friendly tool with GUI that can traverse the assembly graph and bridge its components in real-time as long as the nanopore sequencing process is still running.

We use different hybrid data sets to study the \npgraph{}`s performance and compare with other available tools with similar functionality, such as \npscarf{}, \unicycler{} and \hspades{}. Even though all data were from microbial genomes, the complexity of the input graphs range from few hundred to ten thousands of components, synthesized \emph{in silico} or sequenced from \emph{in vitro} samples.
%Particularly, for the first time, we employ a pipeline in an attempt to complete metagenomics assembly in real-time. 
%We aim to develop a simple approach on top of \npgraph{} to adapt with metagenomics while minimizing additional processing steps. In fact, only one more step for population binning is needed to make \npgraph{} functions on genomes of a mock community.
%Despite a rather straightforward methodology, the pipeline has shown consecutively enhanced assembly from the fragmented short-read DBG graph.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
\subsection*{\npgraph{} -- a tool to resolve assembly graph in real-time}
\npgraph{}'s input consists of Illumina assembly graph resulted from running assembler, e.g. \spades{}~\cite{BankevichNA2012}, $\mathtt{Velvet}$ \cite{Zerbino2008}, $\mathtt{AbySS}$ \cite{Simpson2009} on Illumina short reads, together with long reads from third generation sequencing technology (Oxford Nanopore Technology, Pacbio).
The long reads will be aligned with the contigs in the assembly graph to indicate longer paths that should be traversed. These local paths, given sufficient data, are expected to untangle the complicated graph and guide to the global Eulerian paths (or cycles if possible) that represent the entire genomic sequences. 
\npgraph{} can be invoked and fully function from the command-line interface. In addition, in order to aid the visualization of the assembly process, a GUI has been developed as well.

@@

The GUI includes the dashboard for control the settings of the program and another pop-up window for a simple visualization of the assembly graph in real-time (Figure~\ref{figure:npgraph_gui}).
In this interface, the assembly graph loading stage is separated from the actual assembly process so that users can check for the graph quality first before carry out any further tasks. The box numbered \textbf{1} on Figure~\ref{figure:npgraph_gui} is designed for this task.
Only after an assembly graph is loaded successfully, users can move to box \textbf{2} to specify the nanopore input data.
Settings for an aligner (\bwa{} or \minimap{}) in box \textbf{3} is required if the input is the raw sequences in FASTA/FASTQ format. Another option is to run the alignment independently and provide SAM/BAM input for the next stage of bridging and assembly. This stage is controlled by buttons in box \textbf{4}: the START button ignites the process while the STOP button can prematurely terminate it and output the assembly result till that moment. The plots from the right panel (\textbf{5}) depicts real-time statistics of the assembly contigs inferred from the graph.
From the second window (\textbf{6}), the colored vertices imply unique contigs while the white ones involve either unspecified or repetitive elements. The number of different colors (other than white) indicates the amount of abundant groups being detected as population bins (\EG{} chromosome versus different plasmids, or different bins in metagenomics).

A proper combination of command line and GUI can provide an useful streaming pipeline that copes well with MinION output data. The practice is similar to the previous developed pipelines~\cite{CaoGC2016,Cao2017scaffolding,Nguyen2017barcode} that allow the analysis to take place abreast to a nanopore sequencing run.

\subsection*{Evaluation using synthetic data}
To evaluate the performance of the method, \npgraph{} was tested along with \spades{}, its hybrid assembly module~\cite{AntipovKM2015}, \npscarf{}, and Unicycler version 0.4.6 using Unicycler`s synthetic data set~ \cite{Wick2017unicycler} . 
The data set is a simulation of Illumina and MinION raw data, generated \emph{in silico} based on random and available microbial references. 
We ran all hybrid assembly methods in batch-mode and the reciprocal results were examined by QUAST 5.0.2~\cite{Mikheenko2018quast5}.

@@

Table~\ref{table:npgraph_compare} shows comparative results running different methods on 5 synthetic data sets, simulated from complete genomes of \emph{Mycobacterium~tuberculosis} H37Rv, \emph{Klebsiella~pneumoniae} 30660/NJST258\_1, \emph{Saccharomyces~cerevisiae} S288c, \emph{Shigella~sonnei} 53G and \emph{Shigella~dysenteriae} Sd197.
In the first column of applied methods, beside \unicycler{} and $\mathtt{hybridSPAdes}$, the preceding real-time scaffolder \npscarf{} was included as well.
On the other hand, 2 settings to run \npgraph{} were considered for comparison.

To align the long reads to the assembly graph components, either \bwa{}~\cite{Li2013} or \minimap{}~\cite{Li2016} could be invoked from \npgraph{}.
While \bwa{} is a well-known classic aligner that can be adapted to work with third generation sequencing data, \minimap{} is the state-of-the-art specially designed for this data type. 
In \npgraph{}, the former option was inherited from \npscarf{} pipeline with identical parameters for the alignment algorithm
(-$\mathtt{k}$11~-$\mathtt{W}$20~-$\mathtt{r}$10~-$\mathtt{A}$1~-$\mathtt{B}$1~-$\mathtt{O}$1~-$\mathtt{E}$1~-$\mathtt{L}$0~-$\mathtt{a}$~-$\mathtt{Y}$)
while the latter was used with the sensitive settings (-$\mathtt{k}$15~-$\mathtt{w}$5) recommended for MinION data.
For this synthetic data, \bwa{} reported more but shorter hits than \minimap{} thus was responsible for higher number of false positive alignments.
As a result, bridging paths induced by using \bwa{} was slightly less accurate due to more noises from the smaller \emph{steps} inbetween (Figure~\ref{figure:npgraph_merging}).
From all data sets in Table~\ref{table:npgraph_compare}, we can observe higher errors, which are the sum of mismatches and indels, per 100 Kbp from \npgraph{} using the classic tool.
However, under almost circumstances, using either aligners would result in final assemblies with comparable qualities.
This is due to the fact that false possitive alignments mostly come from shorter \emph{stepping} nodes instead of longer matches from the \emph{anchors}.
Also, in reality, the error profile of sequencing data should be varied, not following the fixed pattern from this \emph{in silico} simulation and \bwa{} might indeed perform better sometimes. 
But in terms of running time and resources required, \minimap{} proved to be a better option. 
The total CPU hours had been cut down drastically with the new aligner, making \npgraph{} the fastest hybrid assembler available.
This feature is certainly more favoured for a real-time assembly pipeline that normally requires quick responses in very short intervals.
Because \minimap{} is supposed to replace \bwa{} for long-read sequencing data alignment, it likewise being chosen as the default aligner in the \npgraph{} pipeline.

Overall, as expected from hybrid methods, the error rate of the draft assembly hardly exceeded 100 bp per 100 Kbp (equivalent to 0.1\%) in almost every cases.
The only exception is \npscarf{} as it would sacrify the accuracy for the continuity by using the nanopore bases to fill in the gap between 2 unlinked unique contigs. 
This happened for 2 \emph{Shigella} simulated genomes containing unusually long stretches of repetitive elements ($> 10$Kbp) for bacteria strains. Especially for the last case for \emph{Shigella dysenteriae}, \npscarf{} failed to finish its exhaustive path finding within reasonable time frame thus being excluded from the report.
In fact, the indel errors (typical in nanopore sequencing) were found relatively low in the final contigs (Table~\ref{supp_tab:synthetic_benchmark}).
The majority of the differences accounted for the mismatched nucleotides caused by the alternative paths connecting the unique anchors from the backbone of the assembly.
This phenomenon may root from homologous repeats or sequencing errors of the genome.

Amongst all assemblers, \unicycler{} applies an algorithm based on semi-global (or glocal) alignments~\cite{Brudno2003glocal} with the consensus long reads using $\mathtt{SeqAn}$ library. When testing with these data sets, it was on top in terms of computational cost but returned the most reliable and complete assembly.
$\mathtt{hybridSPAdes}$ reported decent results with high fidelity at base level. As the trade-off, there were fewer connections satisfying its quality threshold, resulting in the fragmented assemblies with lower N50 compared to the other hybrid assemblers. This behaviour was clearly reflected in the last, also the most challenging task of assembly \emph{S.~dysenteriae}.

Regarding streaming algorithms, \npscarf{}, as aforementioned, used the most greedy scaffolding approach that might introduce extra mis-assemblies and errors.
They might be negligible for the majority of common bacterial genomes with modest repeat profiles, but in extreme cases ,\EG{} \emph{Shigella} data sets, the obtained results would be more error-proned necessitating extra post-processing, polishing steps afterward.
On the other hand, \npgraph{} significantly reduced the errors compared to \npscarf{}, sometimes even be the best option \EG{} for \emph{M.~tuberculosis} and \kp{}.  
For the yeast \emph{S.~ cerevisiae} data set, its assembly best covered the reference genome but the number of mis-assemblies was up to 5.
The unfavourable figures, namely mis-assemblies and error, were still high in case of \emph{S.~dysenteriae}.
The reason was due to the complicated and extremely fragmented graph components containing a large number of small-scaled contigs that were difficult to be mapped with nanopore data. The progressive path finding module tried to induce the most likely solution from a stream of coarse-grained alignments and blundered in this case.

\subsection*{Hybrid assembly for real data sets}
The complicated graph topology from \emph{S.~dysenteriae} simulation is not common in real-life application. Also, considering the overfitting of \unicycler{} with its own test sets, we collected real-life sample for further comparison.
A number of sequencing data sets from \emph{in vitro} bacterial samples~\cite{George2017M14} were used in this scenario.
The data included both Illumina paired-end reads and MinION sequencing based-call data for each sample.
Due to the unavailability of reference genomes, there were fewer statistics being reported by QUAST for the comparison of the results. 
Instead, we investigated the number of circular sequences and $\mathtt{PlasmidFinder}$ 1.3~\cite{Carattoli2014} mappings to obtain an evaluation on the accuracy and completeness of the assemblies.
Table~\ref{tab:npgraph_real} shows the benchmark results of \npgraph{} (using \minimap{}) against \unicycler{} on three data sets of bacterial species \emph{Citrobacter~freundii}, \emph{Enterobacter~cloacae} and \emph{Klebsiella~oxytoca}.

@@

From the first data set, there was high similarity between final contigs generated by two assemblers.
They shared the same number of circular ultimate sequences, including the chromosomal and other six replicons contigs. 
The only divergence lied on the biggest sequence ($\simeq 5.029$Mbp) when the \unicycler{}'s chromosome was 48 nucleotides longer than that of \npgraph{}.
Five out of six identical replicons were confirmed as plasmids based on the occurences of appropriate Origin of replication sequences (PlasmidFinder database).
In detail, two megaplasmids (longer than $100$Kbp) were classified as IncFIB while the other two mid-size replicons, $85.6$Kbp and $43.6$Kbp, were incL and repA respectively, leaving the shortest one with $2Kbp$ of length as ColRNAI plasmid.
The remaining circular sequence without any hits to the database was $3.2$Kbp long suggesting that it could be phage or newly replicon's DNA.
Lastly, there were still $14.5$Kbp unfinished sequences resulted in 3 linear contigs from \unicycler{} and 2 for \npgraph{} respectively.

The assembly task for \emph{Enterobacter~cloacae} was observed more challenging as the chromosomal DNA sequence not been fully resolved using either method. 
The chromosome size was estimated to be approximately $4.8$Mbp but had been broken into two smaller pieces. 
\npgraph{} returned longer stretches of length $3.324$Mbp and $1.534$Mbp while the figures were $2.829$Mbp and $1.978$Mbp from \unicycler{}'s output.
However, the number of circular sequences detected by \unicycler{} was one more than the other ($2$ versus $1$). They were corresponding to 2 plasmids, namely IncR and repA.
While the latter were recognized by both methods, the longer plasmid sequence was fragmented running \npgraph{}.
Similar to the previous data set, there were around $14$Kbp of data were unable to be finished by the assemblers.

>>>>

\title{
Real-time resolution of short-read assembly graph using ONT long reads}
\shorttitle{Real-time Graph Assembly}

%Authors
\author[1,$\ast$]{Son Hoang Nguyen}
\author[1]{Minh Duc Cao}
\author[1,2,3,4,$\ast$]{Lachlan Coin}

%Affiliation
\affil[1]{Institute for Molecular Bioscience, the University of Queensland, 
St Lucia, Brisbane, QLD 4072 Australia}
\correspondingauthor{
\textsuperscript{$\ast$}
To whom correspondence should be addressed. 
E-mails: l.coin@imb.uq.edu.au,s.hoangnguyen@imb.uq.edu.au
}

\affil[2]{
Department of Microbiology and Immunology, The University of Melbourne, Parkville, 3010, Australia}

\affil[3]{
Department of Clinical Pathology, The University of Melbourne, Parkville, 3010, Australia
}
\affil[4]{Department of Infectious Disease, Imperial College London, London, W2 1NY, UK
}

@@

%Set abstract (see below)
\abstract{
A streaming assembly pipeline utilising real-time Oxford Nanopore Technology (ONT) sequencing data is important for saving sequencing resources and reducing time-to-result. A previous approach implemented in  \npscarf{} provided an efficient streaming algorithm for hybrid assembly but was relatively prone to mis-assemblies compared to other graph-based methods. 
Here we present \npgraph{}, a streaming hybrid assembly tool using the assembly graph instead of the separated pre-assembly contigs. It is able to produce more complete genome assembly by resolving the path finding problem on the assembly graph using long reads as the traversing guide. 
Application to synthetic and real data from bacterial isolate genomes show improved accuracy while still maintaining a low computational cost. 
%On top of that, we apply \npgraph{} on mock community in the very first application to improve metagenomics assembly in real-time with nanopore data. 
\npgraph{} also provides a graphical user interface (GUI) which provides a real-time visualisation of the progress of assembly.
 The tool and source code is available at \url{https://github.com/hsnguyen/assembly}.
} 
%Set keywords
\keywords{hybrid assembly, assembly graph, real-time analysis, nanopore sequencing}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}

%Points:
% Sequencing is now easy and cheap. However time is still the bottlenect, especially for time critical applications

%One particular application is genome assembly

%Highlight the importance of real-time/hybrid assembly

%Gaps: Several tool apptemt to use long reads to

% \begin{itemize}
% \item The importance of having a real-time assembler for bacterial and small genomes: e.g. npscarf 
% \item The difficulty to resolve assembly graph in real-time.
% \item Available hybrid assembly methods using long reads: unicycler, ...
% \item an online algorithm to resolve assembly graph by long reads is needed, ...
% \item Requirements
%     \begin{itemize}
%     \item Support streaming: efficient with high yield (armortized analysis)
%     \item Statistically robust: online modification of the dynamic graph, back tracking edge deletion...
%     \item Standardisation : output GFA in real-time
%     \end{itemize}

Sequencing technology has reached a level of maturity which allows the decoding of virtually any piece of genetic material
which  can be obtained. 
%This provides the potential use of DNA sequencing into a wide range of applications in medical 
%and scientific research. 
However, the time from sample to result remains a barrier to adoption of 
sequencing technology into time critical applications such as infectious disease diagnostics or clinical decision making.  While
there exists real-time sequencing technology such as Oxford Nanopore Technologies (ONT);  algorithms for streaming analyses  of such real-time data are still
in their infancy. Effective streaming methodology will help bridge the gap between potential and practical use.

One particular strength of ONT technology is the production of ultra long reads. This is complementary to the dominant
short read sequencing technology Illumina which is cheaper and has higher per base read quality
but is unable to resolve the complex regions of the genome due to its read length limitation. A natural combination of
the two technologies is in hybrid assembly of genomes. Previously, we developed \npscarf{}~\cite{Cao2017scaffolding} an algorithm to 
scaffold a draft assembly from Illumina sequencing simultaneously with ONT sequencing.  However, \npscarf{} ignores the rich  connectivity information in the short read assembly graph, and as a result is prone to mis-assembly.
%The strategy proved to be
%a time- and cost- effective methodology where practitioners need to generate only sufficient data for completing 
%a draft genome.

 Here, we present \npgraph{}, a novel algorithm to resolve the assembly graph in real-time using long read 
sequencing. \npgraph{} uses the stream of long reads to untangle knots in the assembly graph, which is maintained in memory.
 Because of this, \npgraph{} has better estimation of multiplicity of repeat contigs, resulting in fewer misassemblies. 
 In addition, we develop a visualisation tool for practitioners to monitor
the progress of the assembly process.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}
\subsection*{Resolving assembly graph in real-time}
\npgraph{} makes use of an assembly graph generated from assembling short reads using a de Bruijin
graph method such as \spades{}~\cite{BankevichNA2012}, $\mathtt{Velvet}$~\cite{Zerbino2008}
and $\mathtt{AbySS}$~\cite{Simpson2009}. The assembly graph consists of a list of contigs, and 
possible connections among these contigs. In building the assembly graph, the de Bruijin graph assembler attempts to extend 
each contig as far as possible, until there is more than one possible way of extending due to the repetitive sequences 
beyond the information contained in short reads. Hence each contig has multiple possible connections with others,
 creating knots in the assembly graph. \npgraph{}  uses the connectivity information from long reads to untangle these knots in real-time. 
 With sufficient data, when all the knots are removed, the assembly graph is simplified 
to a path which represents the complete assembly.

\begin{figure}[!hpt]
\centering
\subfloat[Merging bridges]{
    \label{figure:npgraph_merge}
	\includegraphics[width=.48\textwidth]{images/bridge_merging.pdf}
}
\hfill
\subfloat[Knot untangled]{
    \label{figure:npgraph_untangle}
	\includegraphics[width=.48\textwidth]{images/knot.pdf}
}
\caption{Graph resolving algorithm. \ref{figure:npgraph_merge} the bridges suggested by long reads are merged progressively with dynamic programming to find the best path connecting 2 anchors. \ref{figure:npgraph_untangle} A knot (repetitive contig) is unwound following the best path (highlighted in purple) leading to the graph simplification.}
\label{figure:npgraph_pipeline}
\end{figure}

\npgraph{} aligns long reads to the contigs in the assembly graph. When a long read is aligned
to multiple contigs, \npgraph{} constructs candidate paths that are supported by the read. This strategy allows \npgraph{} to progressively update the likelihood of the paths going through a knot.
When sufficient data are obtained, the best path is confidently identified and hence the knot is untangled.
%TODO: two more sentences describing the bridging
In general, the bridging algorithm used to determine the best path is a combination of progressive merging, accumulated scoring and decision making modules.
It operates on each pair of unique contigs, or \emph{anchors}, by using a \emph{bridge} data structure maintaining 2 anchors and a list of \emph{steps} in-between as shown in Figure~\ref{figure:npgraph_merge}. 
A set of candidate paths are listed and the best one can be selected amongst them given enough evidence. 
The de Bruijin graph is subsequently simplified when the bridge is replaced by a single edge representing the best path (\ref{figure:npgraph_untangle}).

%TODO: Because of the converting long reads to contig alignment to mapping to path, npgraph is faster than unicycler, and can be done in real-time

% 
% These local paths, given sufficient data, are expected to untangle the complicated graph and guide to the global Eulerian paths (or cycles if possible) that represent the entire genomic sequences. 
% \npgraph{} can be invoked and fully function from the command-line interface. In addition, in order to aid the visualization of the assembly process, a GUI has been developed as well.

@@

We also provide a Graphical User Interface (GUI) for \npgraph{}. The GUI includes the dashboard for controlling the settings of the program and a window for visualization of the assembly graph 
in real-time (Figure~\ref{figure:npgraph_gui}). 
In this interface, the assembly graph loading stage is separated from the actual assembly process so that users can check for the graph quality first before carry out any further tasks.
A proper combination of command line and GUI can provide an useful streaming pipeline that copes well with MinION output data. The practice is to support the real-time monitoring of
the results from real-time sequencing~\cite{CaoGC2016,Cao2017scaffolding,Nguyen2017barcode} that allow the analysis to take place abreast to a nanopore sequencing run.

% TODO: Extract information from here to caption
% The box numbered \textbf{1} on Figure~\ref{figure:npgraph_gui} is designed for this task.
% Only after an assembly graph is loaded successfully, users can move to box \textbf{2} to specify the nanopore input data.
% Settings for an aligner (\bwa{} or \minimap{}) in box \textbf{3} is required if the input is the raw sequences in FASTA/FASTQ format. Another option is to run the alignment independently and provide SAM/BAM input for the next stage of bridging and assembly. This stage is controlled by buttons in box \textbf{4}: the START button ignites the process while the STOP button can prematurely terminate it and output the assembly result till that moment. The plots from the right panel (\textbf{5}) depicts real-time statistics of the assembly contigs inferred from the graph.
% From the second window (\textbf{6}), the colored vertices imply unique contigs while the white ones involve either unspecified or repetitive elements. The number of different colors (other than white) indicates the amount of abundant groups being detected as population bins (\EG{} chromosome versus different plasmids, or different bins in metagenomics).

\subsection*{Evaluation using synthetic data}
To evaluate the performance of the method, \npgraph{} 1.1 was tested along with \spades{}, \spades{} hybrid from version 3.13.1, ~\cite{AntipovKM2015}, \npscarf{} (japsa 1.7-02), and \unicycler{} version 0.4.6 %%COMMENT_LC - why just indicate unicyler version and not others? 
using \unicycler{}`s synthetic data set~ \cite{Wick2017unicycler} . 
The data set is a simulation of Illumina and MinION raw data, generated \emph{in silico} based on available microbial references.
We ran hybrid assembly methods using the entire nanopore data and the reciprocal results were evaluated by QUAST 5.0.2~\cite{Mikheenko2018quast5}.

@@

Table~\ref{table:npgraph_compare} shows comparative results running different methods on 5 synthetic data sets, simulated from complete genomes of \emph{Mycobacterium~tuberculosis} H37Rv, \emph{Klebsiella~pneumoniae} 30660/NJST258\_1, \emph{Saccharomyces~cerevisiae} S288c, \emph{Shigella~sonnei} 53G and \emph{Shigella~dysenteriae} Sd197.
%FOLLOWING REPEATS WHAT YOU SAID BEFORE
%In the first column of applied methods, beside \unicycler{} and $\mathtt{hybridSPAdes}$, the preceding real-time scaffolder \npscarf{} was included as well.
%On the other hand, 2 settings to run \npgraph{} were studied to have insights into the effect of using different alignment methods on its result.

To align the long reads to the assembly graph components, both \bwa{}~\cite{Li2013} or \minimap{}~\cite{Li2016} were used in conjunction with \npgraph{}. These two methods were chosen due to their proven efficiency and compatibility with streaming data.
While \bwa{} is a well-known classic aligner that can be adapted to work with third generation sequencing data, \minimap{} has been specially designed for this data type. 
%In \npgraph{}, the former's application is similar to \npscarf{} pipeline with identical parameters
%(-$\mathtt{k}$11~-$\mathtt{W}$20~-$\mathtt{r}$10~-$\mathtt{A}$1~-$\mathtt{B}$1~-$\mathtt{O}$1~-$\mathtt{E}$1~-$\mathtt{L}$0~-$\mathtt{a}$~-$\mathtt{Y}$)
%while the latter is used with the sensitive settings 
%(-$\mathtt{k}$15~-$\mathtt{w}$5) 
%recommended for MinION data.
We observed a slightly higher error rate (comprising the sum of mismatches and indels per 100kb) using \bwa{} in comparison to \minimap{} for all simulations in Table~\ref{table:npgraph_compare}.  This is due to the fact that bridging paths induced using \bwa{} were slightly less accurate due to more noise from the smaller \emph{steps} in-between (Figure~\ref{figure:npgraph_merge}).
%For this synthetic data, \bwa{} reported more but shorter hits than \minimap{} thus was responsible for higher number of false positive alignments.
However, under almost circumstances, using either aligner resulted in final assemblies with comparable qualities.
%The reason is that false positive alignments mostly come from shorter \emph{stepping} nodes instead of longer and more important hits for the \emph{anchors} which form the skeleton of the final assembly.
%Also, in reality, the error profile of sequencing data is varied due to many factors, not simply following a pre-defined distribution from an \emph{in silico} simulation.
%In other words, no single setting of a tool is assured the best performance with every data sets.
In terms of running time and resources required, \minimap{} always proved to be a better option, requiring markedly less CPU time than \bwa{}.  Utilising \minimap, \npgraph{} is now the fastest hybrid assembler available.

% Generally for hybrid approaches, the error rate of the draft assembly hardly exceeded 100 bp per 100 Kbp (equivalent to 0.1\%).
% The only exception is \npscarf{} as it would sacrifice the accuracy for the continuity by using the nanopore bases to fill in the gap between 2 unlinked unique contigs. 
% This happened for 2 \emph{Shigella} simulated genomes containing unusually long stretches of repetitive elements ($> 10$Kbp) for bacteria strains. Especially for the last case for \emph{Shigella dysenteriae}, \npscarf{} failed to finish its exhaustive path finding within reasonable time frame thus being excluded from the report.
% In fact, the indel errors (typical in nanopore sequencing) were found relatively low in the final contigs (Table~\ref{supp_tab:synthetic_benchmark}).
% The majority of the differences accounted for the mismatched nucleotides caused by the alternative paths connecting the unique anchors from the backbone of the assembly.
% This phenomenon may root from homologous repeats or sequencing errors of the genome.

Amongst all assemblers, \unicycler{} applies an algorithm based on semi-global (or glocal) alignments~\cite{Brudno2003glocal} with the consensus long reads generated with the $\mathtt{SeqAn}$ library. With all of the data sets tested, \unicycler required the most computational resources, but it also returned  fewer mis-assemblies than the other approaches with a comparable rate of error (indels and mismatches) to \npgraph{}.
$\mathtt{hybridSPAdes}$ reported decent results with high fidelity at base level. As the trade-off, there were fewer connections satisfying its quality threshold, resulting in the fragmented assemblies with lower N50 compared to the other hybrid assemblers. This behaviour was clearly reflected in the last, also the most challenging task of assembly \emph{S.~dysenteriae}.

Of the two streaming algorithms, \npscarf{} utilizes a fast but greedy scaffolding approach that can lead to mis-assemblies and errors.
For bacterial genomes with modest complexity these are minimal (e.g. \kp), but for those with severe repetitive elements, extra calibrations are needed to prevent the mis-assembly due to ambiguous alignments.
On the other hand, \npgraph{} significantly reduced the errors compared to \npscarf{}, sometimes even proved to be the best option \EG{} for \emph{M.~tuberculosis} and \kp{}.  
For the yeast \emph{S.~cerevisiae} data set, the \npgraph{} assembly best covered the reference genome but the number of mis-assemblies was up to 5.
The unfavourable figures, namely mis-assemblies and error, were still high in case of \emph{S.~dysenteriae}, 
 due to the complicated and extremely fragmented graph components containing a large number of small-scaled contigs that were difficult to map with nanopore data. The progressive path finding module tried to induce the most likely solution from a stream of coarse-grained alignments, without fully succeeding.

\subsection*{Hybrid assembly for real data sets}
%The complicated graph topology from \emph{S.~dysenteriae} simulation is not common in real-life application.  %%CAN WE SUPPORT BETTER
%Also, considering the potential for performance bias of \unicycler{} on its own test sets, we collected real-life sample for further comparison of \unicycler and \npgraph.
A number of sequencing data sets from \emph{in vitro} bacterial samples~\cite{George2017M14} were used to further explore differences in performance between \npgraph{} and \unicycler{}.
The data included both Illumina paired-end reads and MinION sequencing based-call data for each sample.
Due to the unavailability of reference genomes, there were fewer statistics reported by QUAST for the comparison of the results. 
Instead, we investigated the number of circular sequences and $\mathtt{PlasmidFinder}$ 1.3~\cite{Carattoli2014} mappings to obtain an evaluation on the accuracy and completeness of the assemblies (
Table~\ref{tab:npgraph_real}) 
%shows the benchmark results of \npgraph{} (using \minimap{}) against \unicycler{} 
on three data sets of bacterial species \emph{Citrobacter~freundii}, \emph{Enterobacter~cloacae} and \emph{Klebsiella~oxytoca}.

@@

There was high similarity between final contigs generated by two assemblers on all of these datasets.
For the \emph{Citrobacter freundii}  dataset, they share the same number of circular  sequences, including the chromosomal and other six replicons contigs in the , with only 48 nucleotides  difference in the length of the main chromosome. 
Five out of six identical replicons could be confirmed as plasmids based on the occurence of  origin of replication sequences from  the PlasmidFinder database.
In detail, two megaplasmids (longer than $100$Kbp) were classified as IncFIB while the other two mid-size replicons, $85.6$Kbp and $43.6$Kbp, were incL and repA respectively, leaving the shortest one with $2Kbp$ of length as ColRNAI plasmid.
The remaining circular sequence without any hits to the database was $3.2$Kbp long suggesting that it could be phage or a cryptic plasmid. 
Both assemblers had $14.5$Kbp of unfinished sequences split amongst 3 linear contigs from \unicycler{} and 2 for \npgraph{}.

The assembly task for \emph{Enterobacter~cloacae} was more challenging and the chromosomal DNA remained fragmented in two contigs for both methods ( of length $3.324$Mbp and $1.534$Mbp for \npgraph{} compared to $2.829$Mbp and $1.978$Mbp for \unicycler{}). Both methods detected two plasmids ( IncR and repA), and \unicycler{} returned comlpete circular sequences for both plasmids, while \npgraph{} returned circular sequence for one plasmid, while the other was fragmented into two contigs. Similar to the assembly of \emph{Citrobacter freundii} , there was around $14$Kbp of data which was unable to be finished by the assemblers (split into 2 and 4 contigs for \unicycler{} and \npgraph{} respectively).
